\section{Choice of Algorithms}
\label{methodology_algorithmchoice}
A search-based approach was chosen in order to solve the problem of generating balanced StarCraft II maps. The method of choice for performing search-based procedural content generation is for the most part evolutionary computation\cite{togelius2011search}(genetic search). Evolutionary computation seeks to optimize the fitness (quality) of solutions in order to find good solutions. The quality of a StarCraft II map consists of multiple parameters, such as distance between bases, safety of resources, number of choke points, etc. Because it is not always desirable to add the fitness of these parameters together into a single fitness function value, multi-objective evolutionary algorithms are sometimes used in order to find the optimal compromises between the different parameters. In contrast, the novelty search paradigm seeks to optimize novelty (divergence from past discoveries), which leads to many (often radically) different solutions in the search space. Because it is not reliant on a fitness function, novelty search is effective when dealing with problems where creating an effective fitness function is difficult, such as when the problem is deceptive or difficult to quantify.

Both evolutionary computation and novelty search faces some significant challenges when the search space is constrained\cite{liapis2014constrained}, such as when generating StarCraft II maps. Many solutions are infeasible due to not fulfilling one or more of the strict constraints that good StarCraft II maps impose, e.g. a ground path between starting bases. In evolutionary computing, a common method for dealing with constraints is to use penalty scores, while \textit{minimal criteria novelty search} is an extension of novelty search that can handle constraints.

As part of this thesis, a standard genetic algorithm, a multi-objective evolutionary algorithm, and constrained novelty search have been implemented in order to test and compare their runtime, and the quality and novelty of the maps they produce.